---
title: "Plot a MI network"
output:
  html_document:
    df_print: paged
---

This is an extract of the scripts in step 4. The idea is to illustrate graph construction from one of the functions over-represented in the SGCCA

## 1. Get a matrix with all the features in the SGCCA components where the function is over-represented

Run in the same folder than subtype.eigeNormi & the enrichment results

Rscript get_matrix <GO:0007156> Basal

```{r}
suppressPackageStartupMessages({library(tidyverse)})

fun="GO:0007156"#args[1]
subty="Basal"#args[2]

#get the components linked to the function
if(length(grep("GO",fun))>0){
 enrich=read_tsv("BP.enrichment",show_col_types = FALSE)
}else{
 enrich=read_tsv("KEGG.enrichment",show_col_types = FALSE)	
}
comp=enrich%>%filter(ID==fun&subtype==subty)%>%
			  dplyr::select(component)%>%unlist
head(comp)
```

```{r}
#get the features selected in those components
selected=read_tsv(paste(subty,"stable",sep='.'),,show_col_types = FALSE)
features=selected%>%filter(component%in%comp)%>%
		 distinct(variable)%>%unlist
#print(paste("Function has",length(features),"features associated",sep=' '))
head(features)
```

```{r}
#get the data
data=data.table::fread(paste(subty,"eigeNormi",sep='.'))
data=data[data$V1%in%features,]
#write_tsv(data,paste(fun,subty,"mtrx",sep='.'))#needed for puma
data[1:3,1:3]
```

## 2. Run ARACNE with obtained matrix

```{bash}
mv GO\:0007156.Basal.mtrx ARACNE-multicore-master/launch/
cd ARACNE-multicore-master/launch/
bash run.sh GO\:0007156.Basal.mtrx
head GO\:0007156.Basal.sort
mv GO\:0007156.Basal.sort ../../
cd ../../
```

If something fails probably is due to an incorrect parsing of the parameters to join, they must be: matrix name, ARACNE kernel width & column index name

## 3. Keep just the edges with MI above a threshold that is specific for the edge type

Rscript MIfilter.R <GO:0007156.Basal.sort> → <GO:0007156.Basal.filtered.alt>

If ARACNE MI distributions are not significantly different MIfilter.R outuputs file.filtered & file.filtered.alt.

file.filtered.alt uses a single threshold no matter the edge type

```{r}
edges=read_tsv("GO:0007156.Basal.filtered.alt",show_col_types=F)
head(edges)
```

## 4. Plot the interactions returned by MIfilter.R

Run in the same folder than the results of differential expression/methylation & the file mapping CpG probes to genes

Also Cytoscape has to be already open

Rscript plotGraph.R <GO:0007156.Basal.filtered.alt>

## 5. Get info about the nodes and edges ploted

Cytoscape has to be already open

Rscript annotateGraph.R <GO:0007156.Basal.cys>

```{r}
suppressPackageStartupMessages(library(igraph))
library(RCy3)

net="GO:0007156.Basal.cys"#commandArgs(trailingOnly=TRUE)
id=unlist(strsplit(net,'.',fixed=T))[1]
subty=unlist(strsplit(net,'.',fixed=T))[2]

openSession(net)
g=createIgraphFromNetwork(subty)
chosen=read_tsv("chosen.tsv",show_col_types=F)
head(chosen)
```

```{r}
#get the cluster that contains the function
chosen=chosen%>%filter(ID==id&subtype==subty)
cl=chosen$group
funs=chosen$Description

#get related functions
groups=read_tsv("Groups_per_component.tsv",show_col_types=F)
groups=groups%>%filter(subtype==subty,group==cl)
if(nrow(groups)>1){
	funs=groups%>%select(Description)%>%unlist}
funs
```

```{r}
names=data.frame(cbind("name"=V(g)$name,"label"=V(g)$label))

library(biomaRt)
mart=useEnsembl("ensembl",dataset="hsapiens_gene_ensembl",
	version=105)
#https://dec2021.archive.ensembl.org
myannot=getBM(attributes = c("hgnc_symbol","wikigene_description"),
	filter="hgnc_symbol",values=names$label,mart=mart)
head(myannot)
```

```{r}
enriched=list(BP=read_tsv("subsamples/BP.enrichment",show_col_types=F),
	KEGG=read_tsv("subsamples/KEGG.enrichment",show_col_types=F))
known_genes=lapply(enriched,function(x) 
	x%>%filter(Description%in%funs&subtype==subty)%>%
	dplyr::select(Description,geneID)%>%separate_rows(geneID,sep='/'))
head(known_genes$BP)
```

```{r}
#Checking functions in pubmed
library(rentrez)

#search only the nodes that aren't responsible for the enrichment
query=names$label[!names$label%in%known_genes$label]
query=unique(unlist(strsplit(query,",")))
knownfun=lapply(funs,function(x) #for every function in the net
			lapply(query,function(y) #and every non-functional feature
			#search papers connecting them
			entrez_search(db="pubmed",term=paste(x,y,sep=' AND '))))
as.data.frame(do.call(rbind,lapply(knownfun,function(x) do.call(rbind,lapply(x,function(y) cbind(paste(y$ids,collapse=','),y$QueryTranslation))))))[1:2,]
```

```{r}
#Checking edges on pubmed
edges=data.frame(get.edgelist(g))
colnames(edges)[2]="name"
edges=merge(edges,names,by="name")[,2:3]
colnames(edges)[2]="pair1"
colnames(edges)[1]="name"
edges=unique(merge(edges,names,by="name")[,2:3])
colnames(edges)[2]="pair2"
edges=edges%>%separate_rows(pair2,sep=',',convert=T)%>%separate_rows(pair2,sep=',',convert=T)
known=apply(edges,1,function(x) 
	entrez_search(db="pubmed",
				  term=paste(x[1],x[2],sep=' AND ')))
head(known,2)
```
